---
title: "Sachs_GCD_hICP Hybrid"
author: "Hongxu Zhu and James Long"
date: "5/23/2022"
output:
  html_document:
    theme: united
    toc: yes
    toc_collapsed: yes
    toc_float: yes
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(kableExtra)
library(AER)
library(InvariantCausalPrediction)
library(ggplot2)
library(readxl)
```

# Summary

* Code is adapted from Sachs_GCD_hICP.Rmd
* Removed non-centered model fits
* Included argument for fitting hybrid estimator (see `hybrid <- TRUE`)
    * Probably best to compare CD with hybrid in real data section because hybrid can directly leverage fact that some variables have change in mean but not variance
    
    
## TODO

* Need to Update GCD graphical model because now using hybrid fit
* Eliminate non-standardized fits
    * Currently GCD is being fit with both standardized and non-standardized (variance set to 1) predictors. Was there any difference in fit quality? Probably should only include standardized
* Suggestion for Table for multivariate analysis:
    * Columns for CD and Hybrid Estimator and one rows for each marker other than PIP2. Entries of table will have cofficient with 95\% CI for effect of PIP2 on marker. hiddenICP will not find any effect (which is clearly wrong based on our univariate analysis) but hybrid estimator will find several significant effects.
* Once code is complete, can move this and univariate analysis into iv-cd-comp folder

# Read / Setup Data

```{r}
## in folder iv-cd-comp
## functions for fitting the GCD and hybrid estimator

## fits gcd given kronecker product and weight matrix
GCD_fit <- function(XkrE,X,y,What){
  A <- t(X)%*%XkrE%*%What%*%t(XkrE)%*%X
  B <- colSums(t(t(X)%*%XkrE%*%What%*%t(XkrE))*y)
  solve(A,B)
}

## computes rowwise kronecker product of A and B
RowwiseKron <- function(A,B){
  return(A[,rep(seq(ncol(A)),each = ncol(B)),drop=FALSE]*B[,rep(seq(ncol(B)),ncol(A)),drop=FALSE])
}

## fits GCD or hybrid (CD + IV constraints) estimator
##
## note: in overidentified case (where number of constraints
## exceeds number of predictors), the estimator implements a
## two step algorithm for estimating the optimal weight matrix.
## this two step estimator uses a two stage least squares like
## initial estimate of the weight matrix W
##
## returns point estimates and covariance matrix
##
##
## arguments
##       E : n x q matrix containing instruments/environments
##       X : n x p matrix of predictors
##       y : length n vector with response
##  center : should y and X columns be set to mean 0
##  hybrid : should instrumental variable constraints be added to estimator
##
##  note on coding E: if using r environments 0, 1, . . ., r-1
##                    column j of E should be 1 if
##                    observation is environment j, 0 o.w.
GCD <- function(E,X,y,center=TRUE,hybrid=FALSE){
  ## normalize data
  n <- nrow(E)
  E <- t(t(E) - colMeans(E))
  if(center){
    X <- t(t(X) - colMeans(X))
    y <- y - mean(y)
  }
  ## compute X dot E (kronecker product of X and E)
  XkrE <- RowwiseKron(X,E)
  ## for hybrid estimator, IV orthogonality constraints are added
  if(hybrid){
    XkrE <- cbind(XkrE,E)
  }
  ## initial weight matrix, motivated by tsls
  What <- solve((1/n)*t(XkrE)%*%XkrE)
  ## in overidentified case, need two steps
  if(ncol(XkrE)>ncol(X)){
    betahat <- GCD_fit(XkrE,X,y,What)
    deltahat <- y - colSums(t(X)*betahat)
    temp <- XkrE*deltahat
    What <- solve((1/n)*t(temp)%*%temp)
  }
  ## compute estimate and estimate of asymptotic variance
  betahat <- GCD_fit(XkrE,X,y,What)
  deltahat <- y - colSums(t(X)*betahat)
  temp <- XkrE*deltahat
  V <- t(temp)%*%temp/n
  M <- -(1/n)*t(XkrE)%*%X
  betahat_var <- solve(t(M)%*%solve(V)%*%M)/n
  return(list(coefficients=betahat,
              cov=betahat_var))
}
```


```{r, include=FALSE}
## import the original dataset

## Separate Dataset
##folder <- "C:/Users/User/Downloads/sachs/"
folder <- "data/"
obs1 <- read_excel(paste0(folder,"1. cd3cd28.xls"))
obs2 <- read_excel(paste0(folder,"2. cd3cd28icam2.xls"))
obs <- rbind(obs1,obs2)
con1 <- read_excel(paste0(folder,"3. cd3cd28+aktinhib.xls"))
con2 <- read_excel(paste0(folder,"4. cd3cd28+g0076.xls"))
con3 <- read_excel(paste0(folder,"5. cd3cd28+psitect.xls"))
con4 <- read_excel(paste0(folder,"6. cd3cd28+u0126.xls"))
con5 <- read_excel(paste0(folder,"7. cd3cd28+ly.xls"))
con6 <- read_excel(paste0(folder,"8. pma.xls"))
con7 <- read_excel(paste0(folder,"9. b2camp.xls"))

names(obs) <- c("Raf","Mek","Plcg","PIP2","PIP3","Erk","Akt","PKA","PKC","P38","Jnk")
names(obs1) <- c("Raf","Mek","Plcg","PIP2","PIP3","Erk","Akt","PKA","PKC","P38","Jnk")
names(con1) <- c("Raf","Mek","Plcg","PIP2","PIP3","Erk","Akt","PKA","PKC","P38","Jnk")
names(con2) <- c("Raf","Mek","Plcg","PIP2","PIP3","Erk","Akt","PKA","PKC","P38","Jnk")
names(con3) <- c("Raf","Mek","Plcg","PIP2","PIP3","Erk","Akt","PKA","PKC","P38","Jnk")
names(con4) <- c("Raf","Mek","Plcg","PIP2","PIP3","Erk","Akt","PKA","PKC","P38","Jnk")
names(con5) <- c("Raf","Mek","Plcg","PIP2","PIP3","Erk","Akt","PKA","PKC","P38","Jnk")
names(con6) <- c("Raf","Mek","Plcg","PIP2","PIP3","Erk","Akt","PKA","PKC","P38","Jnk")
names(con7) <- c("Raf","Mek","Plcg","PIP2","PIP3","Erk","Akt","PKA","PKC","P38","Jnk")

obs1t <- asinh(obs1)
con1t <- asinh(con1)
con2t <- asinh(con2)
con3t <- asinh(con3)
con4t <- asinh(con4)
con5t <- asinh(con5)
con6t <- asinh(con6)
con7t <- asinh(con7)
```

```{r, include=FALSE}
## combine dataset to fit GCD, including 4 conditions

sachs_GCD <- data.frame(rbind(obs1t,con1t,con2t,con3t,con4t))
sachs_Akt <- data.frame(rbind(obs1t,con2t,con3t,con4t))
sachs_PKC <- data.frame(rbind(obs1t,con1t,con3t,con4t))
sachs_PIP2 <- data.frame(rbind(obs1t,con1t,con2t,con4t))
sachs_Mek <- data.frame(rbind(obs1t,con1t,con2t,con3t))


## code for environment matrix
E1 <- numeric(nrow(sachs_GCD))
E1[(nrow(obs1t)+1):(nrow(obs1t)+nrow(con1t))] <- 1

E2 <- numeric(nrow(sachs_GCD))
E2[(nrow(obs1t)+nrow(con1t)+1):(nrow(obs1t)+nrow(con1t)+nrow(con2t))] <- 1

E3 <- numeric(nrow(sachs_GCD))
E3[(nrow(obs1t)+nrow(con1t)+nrow(con2t)+1):(nrow(obs1t)+nrow(con1t)+nrow(con2t)+nrow(con3t))] <- 1

E4 <- numeric(nrow(sachs_GCD))
E4[(nrow(obs1t)+nrow(con1t)+nrow(con2t)+nrow(con3t)+1): length(E4)] <- 1

E <- as.matrix(cbind(E1,E2,E3,E4))

## intervened dataset
E_Akt <- E[-c((nrow(obs1t)+1):(nrow(obs1t)+nrow(con1t))),-1]
E_PKC <- E[-c((nrow(obs1t)+nrow(con1t)+1):(nrow(obs1t)+nrow(con1t)+nrow(con2t))),-2]
E_PIP2 <- E[-c((nrow(obs1t)+nrow(con1t)+nrow(con2t)+1):(nrow(obs1t)+nrow(con1t)+nrow(con2t)+nrow(con3t))),-3]
E_Mek <- E[-c((nrow(obs1t)+nrow(con1t)+nrow(con2t)+nrow(con3t)+1):nrow(sachs_GCD)),-4]
```


# Helper Functions for GCD

```{r}
##function to get p value and confidence intervals from GCD

GCD_p <- function (Eff, Var, df){
  pvalue <- rep(1,length(Eff))
  pvalue <- pmin(pvalue,2*(1-pt(abs(Eff)/pmax(10^(-10),sqrt(Var)),df=df)))
  return(pvalue)
}

GCD_CI <- function (Eff, Var, df){
  add <- qt(1-0.05/2,df)*sqrt(Var)
  CI <- matrix(nrow=2,ncol=length(Eff))
  CI[1,] <- Eff-add
  CI[2,] <- Eff+add
  CI <- data.frame(CI)
  names(CI) <- names(Eff)
  return(CI)
}
```


# Fit GCD


Should hybrid version of GCD be fit which imposes for IV and CD constraints

```{r}
hybrid <- TRUE
gcd <- FALSE
```

```{r}
## Fit hybrid for each marker

## for Raf
y <- sachs_GCD$Raf
X <- sachs_GCD[,-1]

gcd_Raf <- GCD(E,X,y,center=TRUE,hybrid=hybrid)
Eff <- gcd_Raf$coefficients
Var <- diag(gcd_Raf$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_Raf <- round(GCD_p(Eff,Var,df),digits = 3)
CI_Raf <- round(GCD_CI(Eff,Var,df),digits = 3)

result_Rafh <- data.frame(rbind(p_Raf,round(Eff,3),CI_Raf),row.names = c("P-value","coefficient","CI low", "CI up"))

## for Mek

y <- sachs_Mek$Mek
X <- sachs_Mek[,-2]

gcd_Mek <- GCD(E_Mek,X,y,center=TRUE,hybrid=hybrid)
Eff <- gcd_Mek$coefficients
Var <- diag(gcd_Mek$cov)
df <- nrow(sachs_Mek)-1
K <- ncol(E_Mek)

p_Mek <- round(GCD_p(Eff,Var,df),digits = 3)
CI_Mek <- round(GCD_CI(Eff,Var,df),digits = 3)

result_Mekh <- data.frame(rbind(p_Mek,round(Eff,3),CI_Mek),row.names = c("P-value","coefficient","CI low", "CI up"))


## for Plcg

y <- sachs_GCD$Plcg
X <- sachs_GCD[,-3]

gcd_Plcg <- GCD(E,X,y,center=TRUE,hybrid=hybrid)
Eff <- gcd_Plcg$coefficients
Var <- diag(gcd_Plcg$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_Plcg <- round(GCD_p(Eff,Var,df),digits = 3)
CI_Plcg <- round(GCD_CI(Eff,Var,df),digits = 3)

result_Plcgh <- data.frame(rbind(p_Plcg,round(Eff,3),CI_Plcg),row.names = c("P-value","coefficient","CI low", "CI up"))


## for PIP2

y <- sachs_PIP2$PIP2
X <- sachs_PIP2[,-4]

gcd_PIP2 <- GCD(E_PIP2,X,y,center=TRUE,hybrid=hybrid)
Eff <- gcd_PIP2$coefficients
Var <- diag(gcd_PIP2$cov)
df <- nrow(sachs_PIP2)-1
K <- ncol(E_PIP2)

p_PIP2 <- round(GCD_p(Eff,Var,df),digits = 3)
CI_PIP2 <- round(GCD_CI(Eff,Var,df),digits = 3)

result_PIP2h <- data.frame(rbind(p_PIP2,round(Eff,3),CI_PIP2),row.names = c("P-value","coefficient","CI low", "CI up"))


## for PIP3

y <- sachs_GCD$PIP3
X <- sachs_GCD[,-5]

gcd_PIP3 <- GCD(E,X,y,center=TRUE,hybrid=hybrid)
Eff <- gcd_PIP3$coefficients
Var <- diag(gcd_PIP3$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_PIP3 <- round(GCD_p(Eff,Var,df),digits = 3)
CI_PIP3 <- round(GCD_CI(Eff,Var,df),digits = 3)

result_PIP3h <- data.frame(rbind(p_PIP3,round(Eff,3),CI_PIP3),row.names = c("P-value","coefficient","CI low", "CI up"))

## for Erk

y <- sachs_GCD$Erk
X <- sachs_GCD[,-6]

gcd_Erk <- GCD(E,X,y,center=TRUE,hybrid=hybrid)
Eff <- gcd_Erk$coefficients
Var <- diag(gcd_Erk$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_Erk <- round(GCD_p(Eff,Var,df),digits = 3)
CI_Erk <- round(GCD_CI(Eff,Var,df),digits = 3)

result_Erkh <- data.frame(rbind(p_Erk,round(Eff,3),CI_Erk),row.names = c("P-value","coefficient","CI low", "CI up"))

## for Akt

y <- sachs_Akt$Akt
X <- sachs_Akt[,-7]

gcd_Akt <- GCD(E_Akt,X,y,center=TRUE,hybrid=hybrid)
Eff <- gcd_Akt$coefficients
Var <- diag(gcd_Akt$cov)
df <- nrow(sachs_Akt)-1
K <- ncol(E_Akt)

p_Akt <- round(GCD_p(Eff,Var,df),digits = 3)
CI_Akt <- round(GCD_CI(Eff,Var,df),digits = 3)

result_Akth <- data.frame(rbind(p_Akt,round(Eff,3),CI_Akt),row.names = c("P-value","coefficient","CI low", "CI up"))

## for PKA

y <- sachs_GCD$PKA
X <- sachs_GCD[,-8]

gcd_PKA <- GCD(E,X,y,center=TRUE,hybrid=hybrid)
Eff <- gcd_PKA$coefficients
Var <- diag(gcd_PKA$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_PKA <- round(GCD_p(Eff,Var,df),digits = 3)
CI_PKA <- round(GCD_CI(Eff,Var,df),digits = 3)

result_PKAh <- data.frame(rbind(p_PKA,round(Eff,3),CI_PKA),row.names = c("P-value","coefficient","CI low", "CI up"))

## for PKC

y <- sachs_PKC$PKC
X <- sachs_PKC[,-9]

gcd_PKC <- GCD(E_PKC,X,y,center=TRUE,hybrid=hybrid)
Eff <- gcd_PKC$coefficients
Var <- diag(gcd_PKC$cov)
df <- nrow(sachs_PKC)-1
K <- ncol(E_PKC)

p_PKC <- round(GCD_p(Eff,Var,df),digits = 3)
CI_PKC <- round(GCD_CI(Eff,Var,df),digits = 3)

result_PKCh <- data.frame(rbind(p_PKC,round(Eff,3),CI_PKC),row.names = c("P-value","coefficient","CI low", "CI up"))

## for P38

y <- sachs_GCD$P38
X <- sachs_GCD[,-10]

gcd_P38 <- GCD(E,X,y,center=TRUE,hybrid=hybrid)
Eff <- gcd_P38$coefficients
Var <- diag(gcd_P38$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_P38 <- round(GCD_p(Eff,Var,df),digits = 3)
CI_P38 <- round(GCD_CI(Eff,Var,df),digits = 3)

result_P38h <- data.frame(rbind(p_P38,round(Eff,3),CI_P38),row.names = c("P-value","coefficient","CI low", "CI up"))


## for Jnk

y <- sachs_GCD$Jnk
X <- sachs_GCD[,-11]

gcd_Jnk <- GCD(E,X,y,center=TRUE,hybrid=hybrid)
Eff <- gcd_Jnk$coefficients
Var <- diag(gcd_Jnk$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_Jnk <- round(GCD_p(Eff,Var,df),digits = 3)
CI_Jnk <- round(GCD_CI(Eff,Var,df),digits = 3)

result_Jnkh <- data.frame(rbind(p_Jnk,round(Eff,3),CI_Jnk),row.names = c("P-value","coefficient","CI low", "CI up"))

```

```{r}
## Fit GCD for each marker

## for Raf
y <- sachs_GCD$Raf
X <- sachs_GCD[,-1]

gcd_Raf <- GCD(E,X,y,center=TRUE,hybrid=gcd)
Eff <- gcd_Raf$coefficients
Var <- diag(gcd_Raf$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_Raf <- round(GCD_p(Eff,Var,df),digits = 3)
CI_Raf <- round(GCD_CI(Eff,Var,df),digits = 3)

result_Rafg <- data.frame(rbind(p_Raf,round(Eff,3),CI_Raf),row.names = c("P-value","coefficient","CI low", "CI up"))

## for Mek

y <- sachs_Mek$Mek
X <- sachs_Mek[,-2]

gcd_Mek <- GCD(E_Mek,X,y,center=TRUE,hybrid=gcd)
Eff <- gcd_Mek$coefficients
Var <- diag(gcd_Mek$cov)
df <- nrow(sachs_Mek)-1
K <- ncol(E_Mek)

p_Mek <- round(GCD_p(Eff,Var,df),digits = 3)
CI_Mek <- round(GCD_CI(Eff,Var,df),digits = 3)

result_Mekg <- data.frame(rbind(p_Mek,round(Eff,3),CI_Mek),row.names = c("P-value","coefficient","CI low", "CI up"))


## for Plcg

y <- sachs_GCD$Plcg
X <- sachs_GCD[,-3]

gcd_Plcg <- GCD(E,X,y,center=TRUE,hybrid=gcd)
Eff <- gcd_Plcg$coefficients
Var <- diag(gcd_Plcg$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_Plcg <- round(GCD_p(Eff,Var,df),digits = 3)
CI_Plcg <- round(GCD_CI(Eff,Var,df),digits = 3)

result_Plcgg <- data.frame(rbind(p_Plcg,round(Eff,3),CI_Plcg),row.names = c("P-value","coefficient","CI low", "CI up"))


## for PIP2

y <- sachs_PIP2$PIP2
X <- sachs_PIP2[,-4]

gcd_PIP2 <- GCD(E_PIP2,X,y,center=TRUE,hybrid=gcd)
Eff <- gcd_PIP2$coefficients
Var <- diag(gcd_PIP2$cov)
df <- nrow(sachs_PIP2)-1
K <- ncol(E_PIP2)

p_PIP2 <- round(GCD_p(Eff,Var,df),digits = 3)
CI_PIP2 <- round(GCD_CI(Eff,Var,df),digits = 3)

result_PIP2g <- data.frame(rbind(p_PIP2,round(Eff,3),CI_PIP2),row.names = c("P-value","coefficient","CI low", "CI up"))


## for PIP3

y <- sachs_GCD$PIP3
X <- sachs_GCD[,-5]

gcd_PIP3 <- GCD(E,X,y,center=TRUE,hybrid=gcd)
Eff <- gcd_PIP3$coefficients
Var <- diag(gcd_PIP3$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_PIP3 <- round(GCD_p(Eff,Var,df),digits = 3)
CI_PIP3 <- round(GCD_CI(Eff,Var,df),digits = 3)

result_PIP3g <- data.frame(rbind(p_PIP3,round(Eff,3),CI_PIP3),row.names = c("P-value","coefficient","CI low", "CI up"))

## for Erk

y <- sachs_GCD$Erk
X <- sachs_GCD[,-6]

gcd_Erk <- GCD(E,X,y,center=TRUE,hybrid=gcd)
Eff <- gcd_Erk$coefficients
Var <- diag(gcd_Erk$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_Erk <- round(GCD_p(Eff,Var,df),digits = 3)
CI_Erk <- round(GCD_CI(Eff,Var,df),digits = 3)

result_Erkg <- data.frame(rbind(p_Erk,round(Eff,3),CI_Erk),row.names = c("P-value","coefficient","CI low", "CI up"))

## for Akt

y <- sachs_Akt$Akt
X <- sachs_Akt[,-7]

gcd_Akt <- GCD(E_Akt,X,y,center=TRUE,hybrid=gcd)
Eff <- gcd_Akt$coefficients
Var <- diag(gcd_Akt$cov)
df <- nrow(sachs_Akt)-1
K <- ncol(E_Akt)

p_Akt <- round(GCD_p(Eff,Var,df),digits = 3)
CI_Akt <- round(GCD_CI(Eff,Var,df),digits = 3)

result_Aktg <- data.frame(rbind(p_Akt,round(Eff,3),CI_Akt),row.names = c("P-value","coefficient","CI low", "CI up"))

## for PKA

y <- sachs_GCD$PKA
X <- sachs_GCD[,-8]

gcd_PKA <- GCD(E,X,y,center=TRUE,hybrid=gcd)
Eff <- gcd_PKA$coefficients
Var <- diag(gcd_PKA$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_PKA <- round(GCD_p(Eff,Var,df),digits = 3)
CI_PKA <- round(GCD_CI(Eff,Var,df),digits = 3)

result_PKAg <- data.frame(rbind(p_PKA,round(Eff,3),CI_PKA),row.names = c("P-value","coefficient","CI low", "CI up"))

## for PKC

y <- sachs_PKC$PKC
X <- sachs_PKC[,-9]

gcd_PKC <- GCD(E_PKC,X,y,center=TRUE,hybrid=gcd)
Eff <- gcd_PKC$coefficients
Var <- diag(gcd_PKC$cov)
df <- nrow(sachs_PKC)-1
K <- ncol(E_PKC)

p_PKC <- round(GCD_p(Eff,Var,df),digits = 3)
CI_PKC <- round(GCD_CI(Eff,Var,df),digits = 3)

result_PKCg <- data.frame(rbind(p_PKC,round(Eff,3),CI_PKC),row.names = c("P-value","coefficient","CI low", "CI up"))

## for P38

y <- sachs_GCD$P38
X <- sachs_GCD[,-10]

gcd_P38 <- GCD(E,X,y,center=TRUE,hybrid=gcd)
Eff <- gcd_P38$coefficients
Var <- diag(gcd_P38$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_P38 <- round(GCD_p(Eff,Var,df),digits = 3)
CI_P38 <- round(GCD_CI(Eff,Var,df),digits = 3)

result_P38g <- data.frame(rbind(p_P38,round(Eff,3),CI_P38),row.names = c("P-value","coefficient","CI low", "CI up"))


## for Jnk

y <- sachs_GCD$Jnk
X <- sachs_GCD[,-11]

gcd_Jnk <- GCD(E,X,y,center=TRUE,hybrid=gcd)
Eff <- gcd_Jnk$coefficients
Var <- diag(gcd_Jnk$cov)
df <- nrow(sachs_GCD)-1
K <- ncol(E)

p_Jnk <- round(GCD_p(Eff,Var,df),digits = 3)
CI_Jnk <- round(GCD_CI(Eff,Var,df),digits = 3)

result_Jnkg <- data.frame(rbind(p_Jnk,round(Eff,3),CI_Jnk),row.names = c("P-value","coefficient","CI low", "CI up"))

```



```{r}
##DAG found by GCD with centering

library(DiagrammeR)
grViz("
      digraph dig1{
      graph [overlap = true, fontsize = 10]
      node [shape = circle]
      Raf;Mek;Plcg;PIP2;PIP3;Erk;Akt;PKA;PKC;P38;Jnk
      Mek [style = filled, fillcolor = red]
      PIP2 [style = filled, fillcolor = red]
      Akt [style = filled, fillcolor = red]
      PKC [style = filled, fillcolor = red]
      edge [color = purple]
      {Plcg PIP3 Erk Akt P38 Jnk} -> PKC
      {Mek Plcg PIP2 PIP3 Erk PKA} -> Akt
      {Mek Plcg PIP2 PIP3 Akt PKA PKC P38 Jnk} -> Erk
      {Raf Jnk} -> Mek
      {Mek Erk Akt PKC P38 Jnk} -> Raf
      {Mek PIP2 PIP3 Erk Akt PKA Jnk} -> Plcg
      {Plcg PIP2 PIP3 Erk Akt} ->PKA
      {Mek PIP2 Erk Akt PKA} -> PIP3
      {Plcg PIP3 Erk PKA PKC} -> PIP2
      {Mek PIP2 PIP3 Erk Akt PKA PKC Jnk} -> P38
      {Mek Plcg PIP3 Erk PKA P38} -> Jnk
      }
      [1]: object1
      ")
```



# Fit Hidden ICP

```{r}
library(InvariantCausalPrediction)

E_cd_f <- rep(0,nrow(sachs_GCD))
E_cd_f[(nrow(obs1t)+1):(nrow(obs1t)+nrow(con1t))] <- 1
E_cd_f[(nrow(obs1t)+nrow(con1t)+1):(nrow(obs1t)+nrow(con1t)+nrow(con2t))] <- 2
E_cd_f[(nrow(obs1t)+nrow(con1t)+nrow(con2t)+1):(nrow(obs1t)+nrow(con1t)+nrow(con2t)+nrow(con3t))] <- 3
E_cd_f[(nrow(obs1t)+nrow(con1t)+nrow(con2t)+nrow(con3t)+1): nrow(sachs_GCD)]<-4

E_cd_Akt <- rep(0,nrow(sachs_Akt))
E_cd_Akt[(nrow(obs1t)+1):(nrow(obs1t)+nrow(con2t))] <- 1
E_cd_Akt[(nrow(obs1t)+nrow(con2t)+1):(nrow(obs1t)+nrow(con2t)+nrow(con3t))] <- 2
E_cd_Akt[(nrow(obs1t)+nrow(con2t)+nrow(con3t)+1): nrow(sachs_Akt)]<-3

E_cd_PKC <- rep(0,nrow(sachs_PKC))
E_cd_PKC[(nrow(obs1t)+1):(nrow(obs1t)+nrow(con1t))] <- 1
E_cd_PKC[(nrow(obs1t)+nrow(con1t)+1):(nrow(obs1t)+nrow(con1t)+nrow(con3t))] <- 2
E_cd_PKC[(nrow(obs1t)+nrow(con1t)+nrow(con3t)+1): nrow(sachs_PKC)]<-3

E_cd_PIP2 <- rep(0,nrow(sachs_PIP2))
E_cd_PIP2[(nrow(obs1t)+1):(nrow(obs1t)+nrow(con1t))] <- 1
E_cd_PIP2[(nrow(obs1t)+nrow(con1t)+1):(nrow(obs1t)+nrow(con1t)+nrow(con2t))] <- 2
E_cd_PIP2[(nrow(obs1t)+nrow(con1t)+nrow(con2t)+1): nrow(sachs_PIP2)]<-3

E_cd_Mek <- rep(0,nrow(sachs_Mek))
E_cd_Mek[(nrow(obs1t)+1):(nrow(obs1t)+nrow(con1t))] <- 1
E_cd_Mek[(nrow(obs1t)+nrow(con1t)+1):(nrow(obs1t)+nrow(con1t)+nrow(con2t))] <- 2
E_cd_Mek[(nrow(obs1t)+nrow(con1t)+nrow(con2t)+1): nrow(sachs_Mek)]<-3
```

```{r}
## CD centering

center <- function(x){
  x_c <- x-mean(x)
  return(x_c)
}

## Raf

y <- as.numeric(sachs_GCD$Raf)
X <- as.matrix(sachs_GCD[,-1])

y_c <- center(y)
X_c <- apply(X,2,center)

hICP_Raf <- hiddenICP(X_c,y_c,ExpInd = E_cd_f,alpha=0.1)

## Mek

y <- as.numeric(sachs_Mek$Mek)
X <- as.matrix(sachs_Mek[,-2])

y_c <- center(y)
X_c <- apply(X,2,center)

hICP_Mek <-hiddenICP(X_c,y_c,ExpInd = E_cd_Mek,alpha=0.1)

## Plcg

y <- as.numeric(sachs_GCD$Plcg)
X <- as.matrix(sachs_GCD[,-3])

y_c <- center(y)
X_c <- apply(X,2,center)

hICP_Plcg <-hiddenICP(X_c,y_c,ExpInd = E_cd_f,alpha=0.1)

## PIP2

y <- as.numeric(sachs_PIP2$PIP2)
X <- as.matrix(sachs_PIP2[,-4])

y_c <- center(y)
X_c <- apply(X,2,center)

hICP_PIP2 <-hiddenICP(X_c,y_c,ExpInd = E_cd_PIP2,alpha=0.1)

## PIP3

y <- as.numeric(sachs_GCD$PIP3)
X <- as.matrix(sachs_GCD[,-5])

y_c <- center(y)
X_c <- apply(X,2,center)

hICP_PIP3 <-hiddenICP(X_c,y_c,ExpInd = E_cd_f,alpha=0.1)

## Erk

y <- as.numeric(sachs_GCD$Erk)
X <- as.matrix(sachs_GCD[,-6])

y_c <- center(y)
X_c <- apply(X,2,center)

hICP_Erk <-hiddenICP(X_c,y_c,ExpInd = E_cd_f,alpha=0.1)

## Akt

y <- as.numeric(sachs_Akt$Akt)
X <- as.matrix(sachs_Akt[,-7])

y_c <- center(y)
X_c <- apply(X,2,center)

hICP_Akt <-hiddenICP(X_c,y_c,ExpInd = E_cd_Akt,alpha=0.1)

## PKA

y <- as.numeric(sachs_GCD$PKA)
X <- as.matrix(sachs_GCD[,-8])

y_c <- center(y)
X_c <- apply(X,2,center)

hICP_PKA <-hiddenICP(X_c,y_c,ExpInd = E_cd_f,alpha=0.1)

## PKC

y <- as.numeric(sachs_PKC$PKC)
X <- as.matrix(sachs_PKC[,-9])

y_c <- center(y)
X_c <- apply(X,2,center)

hICP_PKC <-hiddenICP(X_c,y_c,ExpInd = E_cd_PKC,alpha=0.1)

## P38

y <- as.numeric(sachs_GCD$P38)
X <- as.matrix(sachs_GCD[,-10])

y_c <- center(y)
X_c <- apply(X,2,center)

hICP_P38 <-hiddenICP(X_c,y_c,ExpInd = E_cd_f,alpha=0.1)

## Jnk

y <- as.numeric(sachs_GCD$Jnk)
X <- as.matrix(sachs_GCD[,-11])

y_c <- center(y)
X_c <- apply(X,2,center)

hICP_Jnk <-hiddenICP(X_c,y_c,ExpInd = E_cd_f,alpha=0.1)
```

Diagram, blue is found by hybrid, red is found by CD and purple is found by both.

```{r}
library(DiagrammeR)
grViz("
      digraph dig1{
      graph [overlap = true, fontsize = 10]
      node [shape = circle]
      Raf;Mek;Plcg;PIP2;PIP3;Erk;Akt;PKA;PKC;P38;Jnk
      Mek [style = filled, fillcolor = red]
      PIP2 [style = filled, fillcolor = red]
      Akt [style = filled, fillcolor = red]
      PKC [style = filled, fillcolor = red]
      edge [color = blue, minlen = 1]
      {Plcg Erk Akt} -> PKC
      {Mek} -> Akt
      {Mek PKA Jnk} -> Erk
      {PIP3} ->PKA
      {PIP3} -> P38
      PIP2 -> PIP3
      {PIP2 PKA} -> Plcg
      edge [color = red, minlen = 1]
      Akt -> Raf
      {Raf PKC} -> Erk
      PKC -> Jnk
      edge [color = purple, minlen =1]
      Mek -> Raf
      Raf -> Mek
      PIP3 -> PIP2
      {Akt P38} -> Erk
      {Erk PKA} -> Akt
      Akt -> PKA
      P38 -> PKC
      {PKC Jnk} -> P38
      P38 -> Jnk
      }
      [1]: object1
      ")
```

Compare CD, our work vs PNAS result, blue is found by our work, red is from PNAS result, purple is from both

```{r}
grViz("
      digraph dig1{
      graph [overlap = true, fontsize = 10]
      node [shape = circle]
      Raf;Mek;Plcg;PIP2;PIP3;Erk;Akt;PKA;PKC;P38;Jnk
      Mek [style = filled, fillcolor = red]
      PIP2 [style = filled, fillcolor = red]
      Akt [style = filled, fillcolor = red]
      PKC [style = filled, fillcolor = red]
      edge [color = blue, minlen = 1]
      Akt -> Raf
      {Raf PKC P38} -> Erk
      Akt -> PKA
      edge [color = red, minlen =1]
      Plcg -> PIP2
      Jnk -> PKC
      edge [color = purple, minlen=1]
      Mek -> Raf
      Raf -> Mek
      PIP3 -> PIP2
      Akt -> Erk
      {Erk PKA} -> Akt
      P38 -> PKC
      {PKC Jnk} -> P38
      {PKC P38} -> Jnk
      }
      [1]: object1
      ")
```

```{r}
### Plot hybrid vs CD, Plcg
lable1 <- rep("Hybrid", 10)
lable2 <- rep("CD", 10)
coeff <- data.frame(cbind(c(as.numeric(result_Plcgh[2,]),as.numeric(hICP_Plcg$betahat)),c(lable1,lable2)))
CI <- data.frame(cbind(c(result_Plcgh[4,],hICP_Plcg$ConfInt[1,]),c(result_Plcgh[3,],hICP_Plcg$ConfInt[2,]),c(lable1,lable2)))

d1 <- data.frame(cbind(coeff[,1],CI))
x <- rep(row.names(d1[1:10,]),2)
d1 <- data.frame(cbind(x,d1))
names(d1) <- c("name","Coefficient","CI_up","CI_Low","method")
d1$Coefficient=as.numeric(d1$Coefficient)
d1$CI_Low = as.numeric(d1$CI_Low)
d1$CI_up = as.numeric(d1$CI_up)
d1$method = as.character(d1$method)

hybrid_cd_plcg <- d1

ggplot(d1, aes(x=name, y=Coefficient,color=method))+geom_errorbar(width=.1, aes(ymin=CI_Low,ymax=CI_up),position = "dodge")+scale_x_discrete()+ coord_cartesian(ylim = c(-0.5,0.5))+geom_hline(yintercept = 0.2,color="grey",linetype="dashed")+geom_hline(yintercept = -0.2,color="grey",linetype="dashed")+theme(legend.title = element_blank())+theme(axis.title.x = element_blank())
```



```{r}
### Plot hybrid vs CD, PIP3
lable1 <- rep("Hybrid", 10)
lable2 <- rep("CD", 10)
coeff <- data.frame(cbind(c(as.numeric(result_PIP3h[2,]),as.numeric(hICP_PIP3$betahat)),c(lable1,lable2)))
CI <- data.frame(cbind(c(result_PIP3h[4,],hICP_PIP3$ConfInt[1,]),c(result_PIP3h[3,],hICP_PIP3$ConfInt[2,]),c(lable1,lable2)))

d1 <- data.frame(cbind(coeff[,1],CI))
x <- rep(row.names(d1[1:10,]),2)
d1 <- data.frame(cbind(x,d1))
names(d1) <- c("name","Coefficient","CI_up","CI_Low","method")
d1$Coefficient=as.numeric(d1$Coefficient)
d1$CI_Low = as.numeric(d1$CI_Low)
d1$CI_up = as.numeric(d1$CI_up)
d1$method = as.character(d1$method)

hybrid_cd_pip3 <- d1

ggplot(d1, aes(x=name, y=Coefficient,color=method))+geom_errorbar(width=.1, aes(ymin=CI_Low,ymax=CI_up),position = "dodge")+scale_x_discrete()+ coord_cartesian(ylim = c(-0.5,0.5))+geom_hline(yintercept = 0.2,color="grey",linetype="dashed")+geom_hline(yintercept = -0.2,color="grey",linetype="dashed")+theme(legend.title = element_blank())+theme(axis.title.x = element_blank())
```

```{r}
### Plot hybrid vs GCD, Plcg
lable1 <- rep("Hybrid", 10)
lable2 <- rep("GCD", 10)
coeff <- data.frame(cbind(c(as.numeric(result_Plcgh[2,]),as.numeric(result_Plcgg[2,])),c(lable1,lable2)))
CI <- data.frame(cbind(c(result_Plcgh[4,],result_Plcgg[4,]),c(result_Plcgh[3,],result_Plcgg[3,]),c(lable1,lable2)))

d1 <- data.frame(cbind(coeff[,1],CI))
x <- rep(row.names(d1[1:10,]),2)
d1 <- data.frame(cbind(x,d1))
names(d1) <- c("name","Coefficient","CI_up","CI_Low","method")
d1$Coefficient=as.numeric(d1$Coefficient)
d1$CI_Low = as.numeric(d1$CI_Low)
d1$CI_up = as.numeric(d1$CI_up)
d1$method = as.character(d1$method)

hybrid_gcd_plcg <- d1

ggplot(d1, aes(x=name, y=Coefficient,color=method))+geom_errorbar(width=.1, aes(ymin=CI_Low,ymax=CI_up),position = "dodge")+scale_x_discrete()+ coord_cartesian(ylim = c(-0.5,0.5))+geom_hline(yintercept = 0.2,color="grey",linetype="dashed")+geom_hline(yintercept = -0.2,color="grey",linetype="dashed")+theme(legend.title = element_blank())+theme(axis.title.x = element_blank())
```

```{r}
### Plot hybrid vs GCD, PIP3
lable1 <- rep("Hybrid", 10)
lable2 <- rep("GCD", 10)
coeff <- data.frame(cbind(c(as.numeric(result_PIP3h[2,]),as.numeric(result_PIP3g[2,])),c(lable1,lable2)))
CI <- data.frame(cbind(c(result_PIP3h[4,],result_PIP3g[4,]),c(result_PIP3h[3,],result_PIP3g[3,]),c(lable1,lable2)))

d1 <- data.frame(cbind(coeff[,1],CI))
x <- rep(row.names(d1[1:10,]),2)
d1 <- data.frame(cbind(x,d1))
names(d1) <- c("name","Coefficient","CI_up","CI_Low","method")
d1$Coefficient=as.numeric(d1$Coefficient)
d1$CI_Low = as.numeric(d1$CI_Low)
d1$CI_up = as.numeric(d1$CI_up)
d1$method = as.character(d1$method)

hybrid_gcd_pip3 <- d1

ggplot(d1, aes(x=name, y=Coefficient,color=method))+geom_errorbar(width=.1, aes(ymin=CI_Low,ymax=CI_up),position = "dodge")+scale_x_discrete()+ coord_cartesian(ylim = c(-0.5,0.5))+geom_hline(yintercept = 0.2,color="grey",linetype="dashed")+geom_hline(yintercept = -0.2,color="grey",linetype="dashed")+theme(legend.title = element_blank())+theme(axis.title.x = element_blank())
```



## Combined Plot

```{r}
out <- rbind(hybrid_cd_plcg,hybrid_gcd_plcg[hybrid_gcd_plcg$method=="GCD",])

p1 <- ggplot(out, aes(x=name,y=Coefficient,color=method))+
  geom_errorbar(width=.35, aes(ymin=CI_Low,ymax=CI_up),position = "dodge",size=1)+
  scale_x_discrete()+ coord_cartesian(ylim = c(-0.5,0.5))+
  geom_hline(yintercept = 0.2,color="grey",linetype="dashed")+
  geom_hline(yintercept = -0.2,color="grey",linetype="dashed")+
  theme(legend.title = element_blank())+theme(axis.title.x = element_blank()) +
  ylab("Causal Effect on Plcg") +
    theme( # remove the vertical grid lines
           panel.grid.major.x = element_blank() ,
           # explicitly set the horizontal lines (or they will disappear too)
           panel.grid.major.y = element_line( size=.1, color="white" ),
           legend.position = c(0.9, 0.85))
print(p1)


out <- rbind(hybrid_cd_pip3,hybrid_gcd_pip3[hybrid_gcd_pip3$method=="GCD",])


p2 <- ggplot(out, aes(x=name,y=Coefficient,color=method))+
  geom_errorbar(width=.35, aes(ymin=CI_Low,ymax=CI_up),position = "dodge",size=1)+
  scale_x_discrete()+ coord_cartesian(ylim = c(-0.5,0.5))+
  geom_hline(yintercept = 0.2,color="grey",linetype="dashed")+
  geom_hline(yintercept = -0.2,color="grey",linetype="dashed")+
  theme(legend.title = element_blank())+theme(axis.title.x = element_blank()) +
  ylab("Causal Effect on PIP3") +
    theme( # remove the vertical grid lines
           panel.grid.major.x = element_blank() ,
           # explicitly set the horizontal lines (or they will disappear too)
           panel.grid.major.y = element_line( size=.1, color="white" ),
           legend.position = c(0.9, 0.85))
print(p2)





pdf("plcg.pdf",width=6,height=4)
print(p1)
dev.off()

pdf("pip3.pdf",width=6,height=4)
print(p2)
dev.off()



```

# PIP2 Investigation


```{r}
lable <- rep("CD3/CD28", nrow(obs1t))
lable_PIP2 <- rep("CD3/CD28+Psitectorigenin", nrow(con3t))
d2 <- cbind(rbind(obs1t,con3t),condition=c(lable,lable_PIP2))
d2 <- data.frame(d2)


vs <- colnames(d2)
vs <- vs[!(vs %in% c("PIP2","condition"))]

for(ii in 1:length(vs)){
  p <- ggplot(d2, aes_string(x="PIP2", y=vs[ii], color="condition")) +
    geom_point() 
  print(p)
  print(eval(as.name(paste0("result_",vs[ii],"h"))))
}
```

## Akt investigation

```{r}
lable <- rep("CD3/CD28", nrow(obs1t))
lable_Akt <- rep("CD3/CD28+Aktinhibitor", nrow(con1t))
d2 <- cbind(rbind(obs1t,con1t),condition=c(lable,lable_Akt))
d2 <- data.frame(d2)


vs <- colnames(d2)
vs <- vs[!(vs %in% c("Akt","condition"))]

for(ii in 1:length(vs)){
  p <- ggplot(d2, aes_string(x="Akt", y=vs[ii], color="condition")) +
    geom_point() 
  print(p)
  print(eval(as.name(paste0("result_",vs[ii],"h"))))
}
```
